Bayesian Nonparametric System Reliability\\ using Sets of Priors

### 2016-01-31 ijar-npb-ms.tex
Intro section done (hopefully)

\author[ein]{Gero Walter\fnref{fn1}}
\ead{g.m.walter@tue.nl}
\author[oxf]{Louis J.M. Aslett}
\ead{louis.aslett@stats.ox.ac.uk}
\author[dur]{Frank P.A. Coolen}
\ead{frank.coolen@durham.ac.uk}

@@

\fntext[fn1]{Gero Walter was supported by the Dinalog project
``Coordinated Advanced Maintenance and Logistics Planning for the Process Industries'' (CAMPI).}

\begin{abstract}
An imprecise Bayesian nonparametric approach
to system reliability with multiple types of components
is developed.
This allows modelling partial or imperfect prior knowledge on component failure distributions
in a flexible way through bounds on the functioning probability.
Given component level test data these bounds are propagated 
to bounds on the posterior predictive distribution for the reliability of
a new system containing components exchangeable with those 
used in testing.  The method further enables identification 
of prior data conflict at the system level based on component level test data.

There are new results on first-order stochastic dominance for the Beta-Binomial distribution
which make the technique computationally tractable.  These methodological contributions can be immediately used in
applications by reliability practitioners using the freely available
software tools which are provided.
\end{abstract}

\begin{keyword}
System reliability \sep
Survival signature \sep
Imprecise probability \sep
Bayesian Nonparametrics \sep
Prior-data conflict
\end{keyword}
\end{frontmatter}

@@

A critique of the methodological approach to a reliability 
analysis may often encompass a few common concerns.
First, in a parametric 
setting, there may be no particularly strong reason to
believe that the small part of component model
space covered by a particular probability distribution 
necessarily contains the `true' component lifetime distribution.
Further, Bayesian methods may be invoked in order to 
incorporate expert opinion or other knowledge which falls
outside the specific testing data under consideration.  The 
classic concern here is in whether one can truly express 
ones beliefs with the exactness a prior requires.  Finally,
it would be valuable in application to have a means of identifying when the
prior choice is having a strong effect and when not.  Any method 
hoping to address these concerns must do so whilst enabling 
realistic system models (with heterogeneous component types)
and remain computationally tractable.

Herein, we make steps toward addressing these concerns by developing
a nonparametric method which utilises imprecise probability 
to encapsulate more vague or imperfect prior beliefs using
upper and lower probabilities.  This overcomes the concern
about component lifetimes being outside a particular 
parametric family, uses a more flexible prior modelling 
framework and leads to an easy method of detecting prior-data
conflict which should be of concern at the system level.

Furthermore, the method is based on the survival signature, a 
recent development which naturally accommodates 
heterogeneous component types laid out in any arbitrary
manner which maintains coherency of the system.  As will be 
seen in the sequel, the method remains computationally tractable.

\td{*** Some more words on the benefits?}

In Section \ref{sec:survsign} we review the survival 
signature and in Section \ref{sec:nonparamapproach} we 
review the nonparametric approach to Bayesian reliability analysis
upon which our work builds 
\cite{2015:bayessurvsign}.  Section \ref{sec:reparam} details
the reparameterisation of that approach which enables the natural
formulation of the system reliability bounds and which leads to
nice closed form results in some later theory.
Section \ref{sec:setsofbetapriors} lays the ground work to incorporate imprecise probability,
culminating in the main results and contributions of this 
work, detailed in Section \ref{sec:setsofrel}.  Section 
\ref{sec:examples} provides details on the software contributions
of this work and shows two worked examples
demonstrating the practicality and usefulness of the method.

\td{*** LOUIS: Not sure if the above is too compact or ok and whether we need to be more expansive in the orange bit above?  Thoughts appreciated! I've left text below which you may want to draw from? ***}

@@

The paper is organised as follows.
In Section~\ref{sec:survsign}, we describe ***

@@

The posterior predictive distribution for the number of components surviving at time $t$
in a new system, based on the lifetime data and the prior information,
is then a so-called Beta-Binomial distribution,
$C^k_t \mid s^k_t \sim \bebin(m_k, \alpha^k_t + s^k_t, \beta^k_t + n_k - s^k_t)$.
That is, we have
\begin{align*}
P(C^k_t = l_k \mid s^k_t) &= {m_k \choose l_k} \frac{B(l_k + \alpha^k_t + s^k_t, m_k - l_k + \beta^k_t + n_k - s^k_t)}
                                                    {B(\alpha^k_t + s^k_t, \beta^k_t + n_k - s^k_t)} \,,
\end{align*}
where $B(\cdot, \cdot)$ is the Beta function.
This posterior predictive distribution is essentially a Binomial distribution
where the functioning probability $\ptk$ takes values in $[0,1]$
with the probability given by the posterior on $\ptk$.

@@

From the properties of the Beta distribution,
it follows that $\yz = \E[p]$ is the prior expectation for the functioning probability $p$,
and that larger $\nz$ values lead to greater concentration of probability measure around $\yz$,
since $\V(p) = \frac{\yz (1-\yz)}{\nz + 1}$.
Consequently, $\nz$ represents the prior strength and moreover can be directly interpreted as a pseudocount, as will become clear.
%this is clear from the interpretation of $\alpha$ and $\beta$ and \eqref{eq:reparam},
%and will be further illustrated below through revisiting the posterior
%%based on observed lifetimes $\vec{t} = (t_1, \ldots, t_{n})$.
Indeed, consider the posterior given that $s$ out of $n$ components function:
by conjugacy
$p \mid s$ is Beta distributed with updated parameters
\begin{align}
\nn &= \nz + n\,, &
\yn &= \frac{\nz}{\nz + n} \cdot \yz + \frac{n}{\nz + n} \cdot \frac{s}{n}\,.
\label{eq:nyupdate}
\end{align}
Thus, after observing that $s$ out of $n$ components function (at time $t$),
the posterior mean $\yn$ for $p$ is a weighted average of
the prior mean $\yz$ and $s/n$ (the fraction of functioning components in the data),
with weights $\nz$ and $n$, respectively.
Therefore $\nz$ takes on the same role for the prior mean $\yz$
as the sample size $n$ does for the observed mean $s/n$,
leading to the notion of it being a pseudocount.
%Indeed, the higher $\nz$, the higher the weight for $\yz$
%in the weighted average calculation of $\yn$,
%so $\nz$ gives the strength of the prior as compared to the sample size $n$.
%The parameters $\nz$ and $\yz$ have a more intuitive interprtation,

@@

Functions which make it easy to use the methods of this paper have been added to the \textbf{R} package \texttt{ReliabilityTheory} \citep{2015:aslett-RT}.  There are two functions of particular note: \texttt{computeSystemSurvivalSignature} and \texttt{nonParBayesSystemInferencePriorSets}

>>>>

\author[ein]{Gero Walter}
\ead{g.m.walter@tue.nl}
\author[oxf]{Louis J.M. Aslett}
\ead{louis.aslett@stats.ox.ac.uk}
\author[dur]{Frank P.A. Coolen}
\ead{frank.coolen@durham.ac.uk}

@@

\begin{abstract}
An imprecise Bayesian nonparametric approach
to system reliability with multiple types of components
is developed.
This allows modelling partial or imperfect prior knowledge on component failure distributions
in a flexible way through bounds on the functioning probability.
Given component level test data these bounds are propagated 
to bounds on the posterior predictive distribution for the functioning probability of
a new system containing components exchangeable with those 
used in testing.  The method further enables identification 
of prior-data conflict at the system level based on component level test data.
%
New results on first-order stochastic dominance for the Beta-Binomial distribution
make the technique computationally tractable.
Our methodological contributions can be immediately used in
applications by reliability practitioners
as we provide easy to use software tools.
\end{abstract}

\begin{keyword}
System reliability \sep
Survival signature \sep
Imprecise probability \sep
Bayesian nonparametrics \sep
Prior-data conflict
\end{keyword}
\end{frontmatter}

@@

System reliability analysis is concerned with
estimating the lifetime $\Tsys$ of complex systems.
Usually, the goal is to determine the system reliability function $\Rsys(t) = P(\Tsys > t)$
based on the lifetime distributions of system components.

A critique of the methodological approach to a reliability 
analysis may often encompass a few common concerns.
First, in a parametric 
setting, there may be no particularly strong reason to
believe that the small part of component model
space covered by a particular probability distribution 
necessarily contains the `true' component lifetime distribution.
Further, Bayesian methods may be invoked in order to 
incorporate expert opinion or other knowledge which falls
outside the specific testing data under consideration.  The 
classic concern here is in whether one can truly express 
ones beliefs with the exactness a prior distribution requires.  Finally,
it would be valuable in application to have a means of identifying when the
prior choice is having a strong effect and when not.  Any method 
hoping to address these concerns must do so whilst enabling 
realistic system models (with heterogeneous component types)
and remain computationally tractable.

Herein, we make steps toward addressing these concerns by developing
a nonparametric method which utilises imprecise probability \cite{itip,1991:walley}
to model more vague or imperfect prior beliefs using
upper and lower probabilities.  This overcomes the concern
about component lifetimes being outside a particular 
parametric family, uses a more flexible prior modelling 
framework and leads to an easy method of detecting 
conflicts between prior assumptions and observed failure times in test data.
In the general context of Bayesian methods,
this phenomenon is known as \emph{prior-data conflict},
see, e.g., \cite{2006:evans} or \cite{2015:bickel}.
% which should be of concern at the system level.

Furthermore, the method is based on the survival signature \cite{2012:survsign},
a recent development which naturally accommodates 
heterogeneous component types laid out in any arbitrary manner.
%which maintains coherency of the system.
By separating the (time-invariant) system structure from the time-dependent failure probabilities of components,
it allows straightforward and efficient computation of $\Rsys(t)$.
%by partitioning the event $\Tsys > t$ according to the number of functioning components of each type at time $t$.

Our imprecise probability approach provides bounds for $\Rsys(t)$
by computing, for each $t$ in an arbitrarily fine grid of time points ${\cal T}$,
the posterior predictive probability interval for the event $\Tsys > t$. %that the system will function at time $t$.
Assuming the number of functioning components for each type and time $t$ as binomially distributed, 
the intervals are derived from an imprecise Bayesian model using sets of conjugate Beta priors
which allow to specify weak or partial prior information in an intuitive way.
The width of the resulting posterior predictive probability intervals
reflects the precision of the corresponding probability statements:
a short range indicates that the system functioning probability can be quantified quite precisely,
while a large range will indicate that our (probabilistic) knowledge is indeterminate.
In particular, prior-data conflict leads to more cautious probability statements:
When there is not enough data to overrule the prior,
it is unclear whether to put more trust to prior assumptions or to the observations,
and posterior inferences clearly reflect this state of uncertainty by larger ranges.

While the use of imprecise probability methods can often lead to tractability issues,
new results on first-order stochastic dominance for the Beta-Binomial distribution
keep the need for numerical optimization in our model to a minimum.
%make the technique computationally tractable.
%As will be seen in the sequel, the method remains computationally tractable.
%\td{*** Some more words on the benefits?}

In Section \ref{sec:survsign} we review the survival 
signature and in Section \ref{sec:nonparamapproach} we 
review the nonparametric approach to Bayesian reliability analysis
upon which our work builds 
\cite{2015:bayessurvsign}.  Section \ref{sec:reparam} details
the reparameterisation of that approach which enables the natural
formulation of the system reliability bounds, leading to
nice closed form results in some later theory.
Section \ref{sec:setsofbetapriors} lays the ground work to incorporate imprecise probability,
culminating in the main results and contributions of this 
work, detailed in Section \ref{sec:setsofrel}.  Section 
\ref{sec:examples} provides details on the software contributions
of this work and shows two worked examples
demonstrating the practicality and usefulness of the method.

\iffalse
\td{*** LOUIS: Not sure if the above is too compact or ok and whether we need to be more expansive in the orange bit above?  Thoughts appreciated! I've left text below which you may want to draw from? ***}

@@

The paper is organised as follows.
In Section~\ref{sec:survsign}, we describe ***
\fi

@@

The posterior predictive distribution for the number of components surviving at time $t$
in a new system, based on the lifetime data and the prior information,
is then given by a so-called Beta-Binomial distribution,
$C^k_t \mid s^k_t \sim \bebin(m_k, \alpha^k_t + s^k_t, \beta^k_t + n_k - s^k_t)$.
That is, we have
\begin{align*}
P(C^k_t = l_k \mid s^k_t) &= {m_k \choose l_k} \frac{B(l_k + \alpha^k_t + s^k_t, m_k - l_k + \beta^k_t + n_k - s^k_t)}
                                                    {B(\alpha^k_t + s^k_t, \beta^k_t + n_k - s^k_t)} \,,
\end{align*}
where $B(\cdot, \cdot)$ is the Beta function.
This posterior predictive distribution is essentially a Binomial distribution
where the functioning probability $\ptk$ takes values in $[0,1]$
with the probability given by the posterior on $\ptk$.

@@

From the properties of the Beta distribution,
it follows that $\yz = \E[p]$ is the prior expectation for the functioning probability $p$,
and that larger $\nz$ values lead to greater concentration of probability measure around $\yz$,
since $\V(p) = \frac{\yz (1-\yz)}{\nz + 1}$.
Consequently, $\nz$ represents the prior strength and moreover can be directly interpreted as a pseudocount, as will become clear.
%this is clear from the interpretation of $\alpha$ and $\beta$ and \eqref{eq:reparam},
%and will be further illustrated below through revisiting the posterior
%%based on observed lifetimes $\vec{t} = (t_1, \ldots, t_{n})$.
Indeed, consider the posterior given that $s$ out of $n$ components function:
by conjugacy
$p \mid s$ is Beta distributed with updated parameters
\begin{align}
\nn &= \nz + n\,, &
\yn &= \frac{\nz}{\nz + n} \cdot \yz + \frac{n}{\nz + n} \cdot \frac{s}{n}\,.
\label{eq:nyupdate}
\end{align}
Thus, after observing that $s$ out of $n$ components function (at time $t$),
the posterior mean $\yn$ for $p$ is a weighted average of
the prior mean $\yz$ and $s/n$ (the fraction of functioning components in the data),
with weights proportional to $\nz$ and $n$, respectively.
Therefore $\nz$ takes on the same role for the prior mean $\yz$
as the sample size $n$ does for the observed mean $s/n$,
leading to the notion of it being a pseudocount.
%Indeed, the higher $\nz$, the higher the weight for $\yz$
%in the weighted average calculation of $\yn$,
%so $\nz$ gives the strength of the prior as compared to the sample size $n$.
%The parameters $\nz$ and $\yz$ have a more intuitive interprtation,

@@

Functions which make it easy to use the methods of this paper
have been added to the \textbf{R} package \texttt{ReliabilityTheory} \citep{2015:aslett-RT}.
There are two functions of particular note: \texttt{computeSystemSurvivalSignature}
and, implementing the result from Appendix~\ref{ap:proofs} above, \texttt{nonParBayesSystemInferencePriorSets}.
